---
title: "Getting Started with metaTransportEngine"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with metaTransportEngine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`metaTransportEngine` fits aggregate-data meta-analysis transport models for one or more target populations. The package is built around a common study-level model:

- `y_s | theta_s ~ Normal(theta_s, v_s)`
- `theta_s = beta_0 + z_s^T beta + u_s`
- `u_s ~ Normal(0, tau^2)`

For a target modifier profile `z_T`, the transported mean effect is:

- `theta_T = beta_0 + z_T^T beta`

This vignette walks through the frequentist workflow on the bundled example data and then shows how the Bayesian interface fits into the same data contract.

## Load the package and bundled example data

```{r}
library(metaTransportEngine)

dat <- read_example_studies()
targets <- read_example_targets()

moderators <- c("mean_age", "female_pct", "baseline_risk", "followup_months")

head(dat)
targets
```

## Fit the frequentist transport engine

```{r}
fit <- fit_transport_engine(
  dat = dat,
  targets = targets,
  moderators = moderators,
  effect_scale = "log_rr",
  rob_col = "risk_of_bias"
)

fit
```

The core target-level output contains:

- `transported_mean`: expected mean effect in the target population
- `lci` and `uci`: confidence interval for the transported mean
- `lpi` and `upi`: prediction interval for a realized target effect

```{r}
fit$target_estimates
```

You also get overlap checks, extrapolation scores, heterogeneity summaries, and leave-one-out instability summaries.

```{r}
fit$heterogeneity
fit$extrapolation_scores
fit$leave_one_out$summary
```

## Plot transported effects

```{r, fig.width = 8, fig.height = 5, fig.alt = "Forest-style plot showing study estimates and transported target estimates with confidence intervals and prediction intervals."}
plot_transport_forest(fit, effect_label = "log RR")
```

## Export a report bundle

The export helper writes CSV and JSON outputs plus a forest plot image.

```{r}
out_dir <- file.path(tempdir(), "metaTransportEngine-vignette")
save_transport_results(fit, out_dir)
list.files(out_dir)
```

## Bayesian workflow

The Bayesian interface uses the same `dat`, `targets`, and `moderators` inputs and wraps the Stan models bundled in `inst/stan`.

The call looks like this:

```{r, eval = FALSE}
bayes_fit <- fit_bayesian_transport(
  dat = dat,
  targets = targets,
  moderators = moderators,
  effect_scale = "log_rr",
  chains = 2,
  iter = 800,
  warmup = 400,
  seed = 20260404
)

bayes_fit$target_posterior
bayes_fit$target_predictive
```

The Bayesian chunk is not evaluated in the vignette because Stan compilation and sampling are intentionally kept out of routine package builds and CI checks. Use [run_bayesian_demo.R](../scripts/run_bayesian_demo.R) for the full scripted example in the source tree.

## Next steps

From here, the package supports three common workflows:

- use `fit_transport_engine()` when you want a fast, diagnostics-first transport analysis
- use `fit_bayesian_transport()` when you want posterior and predictive target summaries
- use the exported CSV/JSON outputs as inputs for a larger reporting or evidence-synthesis pipeline
